---
title: "Histograms and Barcharts"
author: "William Taylor Bickelmann"
date: 2017-10-09
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(dplyr)
library(ggplot2)
library(ggiraph)
library(Lahman)
```

## Histogram
Visualization

```{r cars}
df<-Master%>%
  select(weight)
  
```

now to make a plot

```{r pressure, echo=FALSE}
ggplot() + #blank slate 
  geom_histogram(data = df, aes(x=weight))
  
```

Now to get rid of the NAs that are screwing with the results

```{r}
df<-df%>%
  filter(!is.na(weight))
```
now lets try again with some better graphics

```{r}
ggplot()+
  geom_histogram(data=df,aes(x=weight),col="black",fill="green") # two argumets were aded; color and fill
```

```{r}
ggplot()+
  geom_histogram(data=df,aes(x=weight),col="black",fill="green", bins = 25) #now to mess with the bins
```

See how the appearance completely changed? You can get more granularity with more bins
```{r}
ggplot()+
  geom_histogram(data=df,aes(x=weight),col="black",fill="green", bins = 50)
```

Well thats pretty much it for histograms as they are pretty simple. 

##Barplots
The prerequisite for barplots are that the Variable of interest is categorical
```{r}
df<-Teams%>%
  filter(yearID==1980)%>%
  select(name,HR)%>%
  arrange(HR)

df$name<-factor(df$name,levels=df$name)

```

Now that we have a categorical variable, lts get on to making a visualization


```{r}
ggplot()+
  geom_bar(data=df, aes(x=name,y= HR), stat="identity",color="white",fill="black")
```
lets try fliping to see if we can make the plot more readable
```{r}
ggplot()+
  geom_bar(data=df, aes(x=name,y= HR), stat="identity",color="white",fill="black") +
  coord_flip()
```


